---
title: "p8105_mtp_rz2570"
author: "Ruilian Zhang"
date: "10/21/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(ggpubr)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


## Problem 1 - Data

```{r read and clean data}
horn_df = read_excel("data/p8105_mtp_data.xlsx", range = "A9:I1230") %>% 
  janitor::clean_names() %>% 
  mutate_at("eop_size_mm", ~replace(., is.na(.), 0)) %>% 
  mutate(sex = factor(sex),
         age_group = factor(age_group),
         eop_size = factor(eop_size),
         eop_visibility_classification = factor(eop_visibility_classification),
         eop_shape = factor(eop_shape),
         fhp_category = factor(fhp_category)
         )

horn_df %>% 
  group_by(sex, age_group) %>% 
  summarize(n = n()) %>%
  pivot_wider(names_from = age_group,
              values_from = n) %>% 
  mutate(sex = recode(sex, '0' = "Female", '1' = "Male")) %>% 
  knitr::kable()
```

* Fill the missing value in  `eop_size_mm` with `0`, factorize categorical variables.  
* `r ncol(horn_df)` related variables from `r nrow(horn_df)` participants form the data. Information includes sex, age,  size, visibility and shape of EOP, size and category of FHP. There are numeric and categorical variables respectively depicting EOP and FHP.  
* The distribution of age is generally the same across sex. 


```{r plot categorical variables}
a = horn_df %>% 
  ggplot(aes(x = eop_size)) +
  geom_bar()

b = horn_df %>% 
  ggplot(aes(x = eop_visibility_classification)) +
  geom_bar()

c = horn_df %>% 
  ggplot(aes(x = eop_shape)) +
  geom_bar()

d = horn_df %>% 
  ggplot(aes(x = fhp_category)) +
  geom_bar()

ggarrange(a, b, c, d)
```

* The `14.6` in `eop_size` and `30.8` in `fhp_category` do not correctly implement the definition based on underlying continuous variable `eop_size_mm` and `fhp_size_mm`.   
* `eop_visibility_classification` correctly implement the definition based on underlying continuous variable `eop_size_mm`.  
* There are some missing values in `eop_shape` which need to be addressed.  
* Further analysis will filter out age group 1 since one of the member is misclassified by age and the other one aged 17, and the study is looking at age 18+.  
* The abnormal values are listed as below:

```{r etract abnormal values}
horn_df %>% 
  filter(age_group == "1") %>% 
  select(sex, age, age_group) %>% 
  knitr::kable()

horn_df %>% 
  filter(eop_size == "14.6") %>%
  select(sex, age, eop_size_mm, eop_size) %>% 
  knitr::kable()

horn_df %>% 
  filter(fhp_category == "30.8") %>% 
  select(sex, age, fhp_size_mm, fhp_category) %>% 
  knitr::kable()
```


## Problem 2 - Visualization

```{r improve original plots}
fhp_plot = horn_df %>% 
  filter(!(age_group == 1)) %>% 
  mutate(age_group = recode(age_group, `2` = "18-29", `3` = "30's", `4` = "40's", `5` = "50's", `6` = "60+", `7` = "60+", `8` = "60+")) %>%
  ggplot(aes(x = age_group, y = fhp_size_mm, fill = sex)) +
  geom_boxplot() +
  labs(x = "Age group (years)",
       y = "Forward Head Protraction Size (mm)") +
  scale_fill_discrete(name = "Sex", labels = c("Female", "Male")) +
  labs(subtitle = "A")

eop_plot = horn_df %>% 
  filter(!(age_group == 1),
         !(eop_size == 14.6)) %>%
  mutate(
    age_group = recode(age_group, `2` = "18-29", `3` = "30's", `4` = "40's", `5` = "50's", `6` = "60+", `7` = "60+", `8` = "60+"),
    eop_enlarged = ifelse(eop_size %in% c(0, 1), "non_enlarged", "enlarged")) %>% 
  group_by(sex, age_group, eop_enlarged) %>% 
  summarize(n_obs = n()) %>% 
  pivot_wider(
    names_from = eop_enlarged,
    values_from = n_obs) %>% 
  mutate(enlarged_ratio = enlarged / (enlarged + non_enlarged)) %>%
  ggplot(aes(x = age_group, y = enlarged_ratio, group = sex, color = sex)) +
  geom_point() +
  geom_line() +
  labs(x = "Age group (years)",
       y = "Ratio of enlarged EOP") +
  scale_color_discrete(name = "Sex", labels = c("Female", "Male")) +
  labs(subtitle = "B")

ggarrange(fhp_plot, eop_plot)
```

* The overall *FHP size* and *EEOP ratio* are generally larger within male than female. There is an increasing trend of FHP size when the age increases, and the highest FHP size is found in 60+ males. There is no clear age trend for EEOP ratio. People aged 18-29 has the highest ratio of EEOP, while people in their 40's have the lowest ratio of EEOP.

```{r association between FHP size and EOP size}
horn_df %>% 
  filter(!(age_group == 1)) %>%
  mutate(
    sex = recode(sex, `0` = "Female", `1` = "Male"), 
    age_group = recode(age_group, `2` = "18-29", `3` = "30's", `4` = "40's", `5` = "50's", `6` = "60+", `7` = "60+", `8` = "60+")) %>%
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm)) + 
  geom_point(color = "blue", size = 1, alpha = .8) +
  facet_grid(sex ~ age_group) +
  labs(subtitle = "C")
```

* Participants aged 18-29 tend to have smaller FHP sizes, but the distribution of *EOP size* against *FHP size* is not significantly different across age and sex groups, and there is no clear association between EOP size and FHP size within each group.  

## Problem 3 – Reproducing reported results

```{r}
count(horn_df, age_group) %>% 
  knitr::kable()

horn_df %>% 
  group_by(sex) %>% 
  summarize(
    n_obs = n(),
    mean_fhp = mean(fhp_size_mm, na.rm = TRUE),
    sd_fhp = sd(fhp_size_mm, na.rm = TRUE),
    se_fhp = sd_fhp ^ 2 / sqrt(n_obs)) %>% 
  knitr::kable()

horn_df %>% 
  filter(!(eop_size == "14.6")) %>% 
  mutate(eeop = ifelse(as.numeric(eop_size) >= 2, 1, 0)) %>% 
  group_by(eeop) %>% 
  summarize(n_obs = n())

horn_df %>% 
  filter(!(age_group == 1)) %>%
  mutate(age_group = recode(age_group, `2` = "18-29", `3` = "30's", `4` = "40's", `5` = "50's", `6` = "60+", `7` = "60+", `8` = "60+")) %>% 
  ggplot(aes(x = age_group, y = fhp_size_mm)) +
  geom_bar(stat = "identity")
  

horn_df %>% 
  select(age, fhp_size_mm) %>% 
  drop_na() %>% 
  filter(age > 60) %>% 
  mutate(fhp_over_40 = ifelse(fhp_size_mm > 40, 1, 0)) %>% 
  group_by(fhp_over_40) %>% 
  summarize(n_obs = n())
```

* The authors’ stated sample sizes in each age group are  inconsistent with the data available.  
* The reported mean for FHP size is consistent with the data available, but standard deviations aren't the same.  
The authors find “the prevalence of EEOP to be 33% of the study population”. What is the definition of EEOP, and what variables can you use to evaluate this claim? Is the finding consistent with the data available to you?
* *EOP* is *enlarged external occipital protuberance*. The claim can be cacluated by `eop size`. Prevalence calculated by this data is 43%, which is different from 33%.  
* FHP is more common in people over 60. Specific value calculated with data is 32.6%.  

## Prblem 4

*Summarize your results, the quality of the data analysis / presentation of results in the original report, and comment on the conclusions of the reports’ authors. Do you think the data provide evidence that cell phones are causing horn growth? What other data would you like to have to address this hypothesis?*

