---
title: "p8105_mtp_rz2570"
author: "Ruilian Zhang"
date: "10/21/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(ggpubr)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


## Problem 1 - Data

```{r read and clean data}
horn_df = read_excel("data/p8105_mtp_data.xlsx", range = "A9:I1230") %>% 
  janitor::clean_names() %>% 
  mutate_at("eop_size_mm", ~replace(., is.na(.), 0)) %>% 
  mutate(sex = factor(sex),
         age_group = factor(age_group),
         eop_size = factor(eop_size),
         eop_visibility_classification = factor(eop_visibility_classification),
         eop_shape = factor(eop_shape),
         fhp_category = factor(fhp_category)
         )

horn_df %>% 
  group_by(sex, age_group) %>% 
  summarize(n = n()) %>%
  pivot_wider(names_from = age_group,
              values_from = n) %>% 
  mutate(sex = recode(sex, '0' = "Female", '1' = "Male")) %>% 
  knitr::kable()
```

* Fill the missing value in  `eop_size_mm` with `0`, factorize categorical variables.  
* `r ncol(horn_df)` related variables from `r nrow(horn_df)` participants form the data. Information includes sex, age,  size, visibility and shape of EOP, size and category of FHP. There are numeric and categorical variables respectively depicting EOP and FHP.  
* The distribution of age is generally the same across sex. 


```{r plot categorical variables}
a = horn_df %>% 
  ggplot(aes(x = eop_size)) +
  geom_bar()

b = horn_df %>% 
  ggplot(aes(x = eop_visibility_classification)) +
  geom_bar()

c = horn_df %>% 
  ggplot(aes(x = eop_shape)) +
  geom_bar()

d = horn_df %>% 
  ggplot(aes(x = fhp_category)) +
  geom_bar()

ggarrange(a, b, c, d)
```

* The `14.6` in `eop_size` and `30.8` in `fhp_category` do not correctly implement the definition based on underlying continuous variable `eop_size_mm` and `fhp_size_mm`.   
* `eop_visibility_classification` correctly implement the definition based on underlying continuous variable `eop_size_mm`.  
* There are some missing values in `eop_shape` which need to be addressed.  
* Further analysis will filter out age group 1 since one of the member is misclassified by age and the other one aged 17, and the study is looking at age 18+.  
* The abnormal values are listed as below:

```{r etract abnormal values}
horn_df %>% 
  filter(age_group == "1") %>% 
  select(sex, age, age_group) %>% 
  knitr::kable()

horn_df %>% 
  filter(eop_size == "14.6") %>%
  select(sex, age, eop_size_mm, eop_size) %>% 
  knitr::kable()

horn_df %>% 
  filter(fhp_category == "30.8") %>% 
  select(sex, age, fhp_size_mm, fhp_category) %>% 
  knitr::kable()
```


## Problem 2 - Visualization

```{r improve original plots}
fhp_plot = horn_df %>% 
  filter(!(age_group == 1)) %>% 
  mutate(age_group = recode(age_group, `2` = "18-29", `3` = "30's", `4` = "40's", `5` = "50's", `6` = "60+", `7` = "60+", `8` = "60+")) %>%
  ggplot(aes(x = age_group, y = fhp_size_mm, fill = sex)) +
  geom_boxplot() +
  labs(x = "Age group (years)",
       y = "Forward Head Protraction Size (mm)") +
  scale_fill_discrete(name = "Sex", labels = c("Female", "Male")) +
  labs(subtitle = "A")

eop_plot = horn_df %>% 
  filter(!(age_group == 1),
         !(eop_size == 14.6)) %>%
  mutate(
    age_group = recode(age_group, `2` = "18-29", `3` = "30's", `4` = "40's", `5` = "50's", `6` = "60+", `7` = "60+", `8` = "60+"),
    eop_enlarged = ifelse(eop_size %in% c(0, 1), "non_enlarged", "enlarged")) %>% 
  group_by(sex, age_group, eop_enlarged) %>% 
  summarize(n_obs = n()) %>% 
  pivot_wider(
    names_from = eop_enlarged,
    values_from = n_obs) %>% 
  mutate(enlarged_ratio = enlarged / (enlarged + non_enlarged)) %>%
  ggplot(aes(x = age_group, y = enlarged_ratio, group = sex, color = sex)) +
  geom_point() +
  geom_line() +
  labs(x = "Age group (years)",
       y = "Ratio of enlarged EOP") +
  scale_color_discrete(name = "Sex", labels = c("Female", "Male")) +
  labs(subtitle = "B")

ggarrange(fhp_plot, eop_plot)
```

* The overall **FHP size** and **EEOP ratio** are generally larger within male than female. There is an increasing trend of FHP size when the age increases, and the highest FHP size is found in 60+ males. There is no clear age trend for EEOP ratio. People aged 18-29 has the highest ratio of EEOP, while people in their 40's have the lowest ratio of EEOP.

```{r association between FHP size and EOP size}
horn_df %>% 
  filter(!(age_group == 1)) %>%
  mutate(
    sex = recode(sex, `0` = "Female", `1` = "Male"), 
    age_group = recode(age_group, `2` = "18-29", `3` = "30's", `4` = "40's", `5` = "50's", `6` = "60+", `7` = "60+", `8` = "60+")) %>%
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm)) + 
  geom_point(color = "blue", size = 1, alpha = .8) +
  facet_grid(sex ~ age_group) +
  labs(subtitle = "C")
```

* Participants aged 18-29 tend to have smaller FHP sizes, but the distribution of **EOP size** against **FHP size** is not significantly different across age and sex groups, and there is no clear association between EOP size and FHP size within each group.  

## Problem 3 – Reproducing reported results

```{r reproduce result}
count(horn_df, age_group) %>% 
  knitr::kable()

horn_df %>% 
  group_by(sex) %>% 
  summarize(
    n_obs = n(),
    mean_fhp = mean(fhp_size_mm, na.rm = TRUE),
    sd_fhp = sd(fhp_size_mm, na.rm = TRUE),
    se_fhp = sd_fhp ^ 2 / sqrt(n_obs)) %>% 
  knitr::kable()

eeop_df = horn_df %>% 
  filter(!(eop_size == "14.6")) %>% 
  mutate(eeop_status = ifelse(as.numeric(eop_size_mm) > 10, 1, 0)) %>% 
  group_by(eeop_status) %>% 
  summarize(n_obs = n())

eeop_prevalence = eeop_df[2,2] / sum(eeop_df[ ,2])

horn_df %>% 
  filter(!(age_group == 1)) %>%
  mutate(age_group = recode(age_group, `2` = "18-29", `3` = "30's", `4` = "40's", `5` = "50's", `6` = "60+", `7` = "60+", `8` = "60+")) %>% 
  ggplot(aes(x = age_group, y = fhp_size_mm)) +
  geom_bar(stat = "identity")
  

fhp_40_df = horn_df %>% 
  select(age, fhp_size_mm) %>% 
  drop_na() %>% 
  filter(age > 60) %>% 
  mutate(fhp_over_40 = ifelse(fhp_size_mm > 40, 1, 0)) %>% 
  group_by(fhp_over_40) %>% 
  summarize(n_obs = n())

fhp_40_proportion = fhp_40_df[2,2] / sum(fhp_40_df[ ,2])
```

* The authors’ stated sample sizes in each age group are  inconsistent with the data available.  
* The reported mean for FHP size is consistent with the data available, but standard deviations aren't the same.  
* **EEOP** is **enlarged external occipital protuberance**. `eop_size` in this data can be used to calculate **EEOP prevalence**, which is `r eeop_prevalence`. This value is slighlty different from author-stated value 33%.  
* FHP is more common in people over 60, which is consistent to the article. Proportion of FHP >40 mm calculated with data is `r fhp_40_proportion`, which is different from the article stated 34.5%.  


## Prblem 4

In the study population, FHP size has a positive association with age. Males are more likely to have larger FHP size and higher EEOP ratio than females. There is no clear association between EOP size and FHP size. The results based on data available is convincible, but I remain doubt on original report since there in inconsistency between author-stated values and real ones. The conclusion made in the report will be more persuasive if the sample size increases and the sampling process is less biased.   
Due to the nature of case study, we cannot draw causal links between EOP and cellphone use. I would like to collect data on average time of electronic device use, collagen synthesis level, habit of poor posture and muscle strength of trapezius from the underlying source population.
